---
title: "Reproducible Research: Peer Assessment 1"
author: "Andrey Vlasenko"
date: "February 19, 2016"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
 
### 1. Setting Working directory

```{r}
rm(list=ls())
WorkingDir <- "C:/Users/vlasenko/YandexDisk/R_Programming/5_CourseProj1/RepData_PeerAssessment1/"
setwd(WorkingDir)
WorkingDir <- getwd()
print(paste("You can find data and results in ",WorkingDir))
```

### 2. Setting file names

```{r}
Url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
F_name <- "repdata_data_activity.zip"
F_name_Data <- "activity.csv"
```

### 3. Downloading and unziping the data (if files aren't exist)

```{r}
if( !(F_name %in% dir()) ) { download.file(Url, destfile = F_name) }
if( !(F_name_Data %in% dir()) ) { unzip(F_name) }
```

### 4. Reading and  the data

```{r}
data <- read.csv(F_name_Data)
library(lubridate)
data$date <- ymd(data$date)
```

## What is mean total number of steps taken per day?

### 1. Evaluating sums of steps for each day (date), mean and median of sums

```{r}
# in our case we have 8 days with all empty values. 
# If we use na.rm then these empty day will stay with 0 sums.
Sum <- aggregate(data$steps, by = list(data$date), sum)
md <- median(Sum$x,na.rm=TRUE)
mn <- mean(Sum$x,na.rm=TRUE)
```

### 2. Plotting histogramm and reporting the mean and the median

```{r}
# histogramm only for not empty days.
hist(Sum$x, main="Histogram of the total number of steps taken each day", 
    xlab = "Steps per day", ylab = "Number of days", labels = TRUE)
text(20000, 25, paste("Median = ",md), col="blue")
text(20000, 20, paste("Mean = ",round(mn,2)),col="green")
abline(v=md, col="blue", lty = 19)
abline(v=mn, col="green", lty = 10)
```
### 3. Median and mean number of steps taken per day:

```{r}
# Median and Mean
print(paste("Median = ",md,"  Mean =", round(mn,5)))
```


## What is the average daily activity pattern?

### 1. Evaluating means of steps for each interval 

```{r}
Sum2 <- aggregate(data$steps, by = list(data$interval), mean, na.rm=TRUE)
md2 <- median(Sum2$x)
mn2 <- mean(Sum2$x)
```

### 2. Plotting graph and reporting the interval with max average number of spets

```{r}
MaxInt <- Sum2[Sum2$x == max(Sum2$x),1]
plot(Sum2$Group.1, Sum2$x, type="l",col="red",main="Daily activity pattern", 
    xlab = "interval", ylab = "steps")
text(1500, 200, paste("Max interval = ",MaxInt),col="black",cex=2)
text(2000, 170, paste("Median = ",round(md2,2)), col="blue")
text(2000, 150, paste("Mean = ",round(mn2,2)),col="green")
abline(h=md2, col="blue", lty = 19)
abline(h=mn2, col="green", lty = 10)
abline(v=MaxInt, col="black", lty = 1)
```


## Imputing missing values

### 1. Reporting missing values

```{r}
library(plyr)
MissedValues <- apply(is.na(data),2,sum)
MissedValues
barplot(MissedValues,main=paste("Missed values (total = ", sum(MissedValues),")"), 
    xlab = "variable", ylab = "Number")
```

### 2. Filling missing values (and verify) using average daily activity pattern

```{r}
dat0 <- mutate(data, CorrectedSteps = ifelse(is.na(steps),
        Sum2[match(interval,Sum2$Group.1),2], steps) )
# Verify the filling of the missing values
missed <- as.integer(names(table(data[is.na(data$steps),3])))
missed_tbl <- cbind(missed,Sum2[match(missed, Sum2$Group.1),])
table(unique(dat0[is.na(dat0$steps),3:4]) == missed_tbl[,2:3])
```

### 3. Evaluating sums of steps for each day (date), mean and median of sums

```{r}
Sum3 <- aggregate(dat0$CorrectedSteps, by = list(dat0$date), sum)
md3 <- median(Sum3$x,na.rm=TRUE)
mn3 <- mean(Sum3$x,na.rm=TRUE)
```

### 4. Plotting histogramm and reporting the mean and the median

```{r}
hist(Sum3$x, main="Histogram of the total number of steps taken each day", 
    xlab = "Steps per day", ylab = "Number of days", labels = TRUE)
text(20000, 25, paste("Median = ",round(md3,2)), col="blue")
text(20000, 20, paste("Mean = ",round(mn3,2)),col="green")
abline(v=md3, col="blue", lty = 19)
abline(v=mn3, col="green", lty = 10)

# Median and Mean
print(paste("Median = ",round(md3,5),"  Mean =", round(mn3,5)))
```

## Are there differences in activity patterns between weekdays and weekends?

### 1. Creating a new factor variable in the dataset with two levels – “weekday” and “weekend”

```{r}
# Setting English for universality
language <- "English"
Sys.setlocale("LC_TIME", language)
dat <- mutate(dat0, day = ifelse( (weekdays(dat0$date,abbreviate = TRUE)=="Sat")
| (weekdays(dat0$date,abbreviate = TRUE)=="Sun") , "weekend", "weekday"))
```

### 2. Evaluating means of steps for each interval

```{r}
Sum4 <- aggregate(dat$CorrectedSteps, by = list(dat$interval, dat$day), mean, na.rm=TRUE)
```

### 3. Making a panel plot containing a time series plots for weekdays and weekends

```{r}
library(ggplot2)
qplot(x=Group.1, y=x, data=Sum4, col=Group.2, facets=Group.2~., geom="line", 
      xlab="Interval",ylab="Number of steps", main = "")
```

```{r}
## For creating reports use
#library(knitr)
#knit2html("PA1_template.Rmd", force_v1 = TRUE)
```
